What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-22 lda-paper.tex
- Motivate why we bother to visualize - [X] Motivate why the visualizations are important - [X] A. state why we visualize, B. state the goal of each visualization In section 4.5 you enumerate several visualizations, but it is again unclear what the goal is. Why do you need these?  What do you intend to detect?  - Clean up intro - [X] intro needs focus on what we did - [X] Sharpen Intro - [X] Make the goal and the achievements more explicit

\author{Abram Hindle and Michael Godfrey and Ric Holt \\
\{ahindle,migod,holt\}@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
University of Waterloo\\
}

@@

Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. We want to see topics that are unique to a
certain time period, displayed across the time axis. Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a single word drawn from its word
distribution.

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer

@@

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking, in practice
there are lot of small trends (90\% of all trends contain 1 topic)
that fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.  The
compact-trend-view is convenient because it shows all of the topic
information, as shown by the zoomed-in view of MaxDB 7.500's
compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
repeating continuous trends easy to pick out, although discontinuous
trends are harder to spot.

The \emph{trend-time-line} (Figure
\ref{fig:trendtimeline}) displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. Therefore if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long.

The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics (approximately $N$ topics multiplied
by $M$ periods), given the allotted space, it is often best to crop
off the trends with only one topic ($1\%$ to $10\%$ of the total
topics), otherwise the tail is long.

>>>>

\author{Abram Hindle, Michael Godfrey and Ric Holt \\
\{ahindle,migod,holt\}@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
University of Waterloo\\
}

@@

\hlb{
We want to see topics that are unique to a
certain time period, displayed across the time axis.
Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. } Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a single word drawn from its
word distribution.  
\hdb{
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
}
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
\hdb{ Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualization provides us with a framework that allows us to
  visually answer many questions about: the spread of topics, how many
  topics were independant per period, what the repeating trends were,
  if the trends dominated or did local topics dominate, how continuous
  trends were, and what the topic text in topic, trend, or period was.
  Since we have multiple questions to answer we have multiple
  visualization, described within this section, that help answer
  different questions.  }

@@

\hlb{ The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
  attempts to show all topics and trends at once across time in a
  compact view that could fit on one page. It tries to give a compact
  summary of the topics and trends across time.  } The
\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
the larger trends to the bottom.  Once the larger trends are stacked
we fill in the gaps with the smaller trends in the same window, and
then stack the smaller trends on top.  Although there is a chance that
gaps will be left due to non-optimal stacking, in practice there are
lot of small trends (90\% of all trends contain 1 topic) that fill in
these gaps quickly.  Different instances of the same trend share the
same color; apart from that, the color of trends is randomly chosen
and color similarity is not meaningful.  The compact-trend-view is
convenient because it shows all of the topic information, as shown by
the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
\ref{fig:zoomedsmear}. The \emph{compact-trend-view} makes repeating
continuous trends easy to pick out, although discontinuous trends are
harder to spot, and provides a general summary of the topics within a project.

\hdb{
The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
attempts to show a summary of trends seperated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
}
The \emph{trend-time-line} displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. Therefore if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long. 
Thus the \emph{trend-time-line} view is used to analyze trends across time.

\hdb{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
  to show a count of how often a trend reoccurs and how many topics
  are related to a trend. It is meant to show the distribution of
  trends by their size in topics and time-span.  } The
\emph{trend-histogram} superficially resembles the trend-time-line.
However, in this view the trends are plotted together by stacking to
the left of their row, thus time information is lost.  The trends are
ordered by the number of topics in the trend.  The trend-histogram
shows the count of instances of a trend and thus indicates which
trends occur the most. Unfortunately due to the large number of topics
(approximately $N$ topics multiplied by $M$ periods), given the
allotted space, it is often best to crop off the trends with only one
topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
long. \hdb{ The \emph{trend-histogram} summarizes the distribution of
trends ordered by size. }

\hdb{ All of these visualization combine to enable an analysis of
  trends and topics. Some views like the \emph{compact-trend-view}
  enable an analysis of local topics while the \emph{trend-histogram}
  and \emph{trend-time-line} focus more on trends. We used these
  visualizations to analyze the database systems that we discuss in
  the following results section.  }
